---
title: "1.5.5 exercises"
author: "Raul Rocha"
date: "2026-02-14"
output: html_document
---

1.
Using ?mpg, the categorical variables are manufacturer, model, trans, drv, fl and class.
When we run mpg, we can see this information by reading the labels under the name of each column (chr, dbl, int).

2. 
```{r}
ggplot(data  = mpg,
       mapping = aes(y = hwy,
                     x =  displ)) +
  geom_point(aes(color = model))
```

```{r}
ggplot(data  = mpg,
       mapping = aes(y = hwy,
                     x =  displ)) +
  geom_point(aes(shape = model))
```

```{r}
ggplot(data  = mpg,
       mapping = aes(y = hwy,
                     x =  displ)) +
  geom_point(aes(color = model, shape =  model))
```
```{r}
ggplot(data  = mpg,
       mapping = aes(y = hwy,
                     x =  displ)) +
  geom_point(aes(color = cty))
```

For a categorical variable with many possible values, color seems to be a more adequate aesthetic than shape. That´s because there are more colors than shapes for R to work with. Also, trying to map a continuous numerical variable to the shape aesthetics raises an error. Mapping continuous numerical variables to the color aesthetics is perfectly reasonable. 

3. 
```{r}
ggplot(data = mpg,
       mapping = aes(y = hwy,
                     x = displ,
                     linewidth = cty)) +
  geom_point()
```

```{r}
ggplot(data = mpg,
       mapping = aes(y = hwy,
                     x = displ,
                     linewidth =hwy)) +
  geom_point()
```

For this particular dataset, since we've tested mapping the aesthetic linewidth to two different variables and they produced exactly the same plot, it appears to have no effect at all.

4. 
```{r}
ggplot(data = mpg,
       mapping = aes(y = hwy,
                     x = displ)) +
  geom_point(aes(color = displ))
```
Mapping the same variable to multiple aesthetics works perfectly fine.

5.
```{r}
ggplot(data = penguins,
       mapping = aes(y= bill_depth_mm,
                     x = bill_length_mm)) +
  geom_point(aes(color = species)) +
  facet_wrap(~species)
```
We can see that species with more bill depth tend to have less bill length.

6.
The following code:
```{r}
ggplot(
  data = penguins,
  mapping = aes(
    x = bill_length_mm, y = bill_depth_mm, 
    color = species, shape = species
  )
) +
  geom_point() +
  labs(color = "Species")
```

yields two different legends because the mapping aes color = species, shape = species provides a legend natively, but the code also added manually a redundant legend by using labs(color = "Species") We can fix this by simply removing the latter:

```{r}
ggplot(
  data = penguins,
  mapping = aes(
    x = bill_length_mm, y = bill_depth_mm, 
    color = species, shape = species
  )
) +
  geom_point()
```


7.

```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill")
ggplot(penguins, aes(x = species, fill = island)) +
  geom_bar(position = "fill")
```
The first plot shows us how the islands are populated by the different penguin species.
The second plot shows us how the different penguin species are distributed between the islands.
